---
title: "Summarise patient characteristics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{a06_summarisePatientCharacteristics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", message = FALSE, warning = FALSE, 
  fig.width = 7
)

library(CDMConnector)
if (Sys.getenv("EUNOMIA_DATA_FOLDER") == "") Sys.setenv("EUNOMIA_DATA_FOLDER" = tempdir())
if (!dir.exists(Sys.getenv("EUNOMIA_DATA_FOLDER"))) dir.create(Sys.getenv("EUNOMIA_DATA_FOLDER"))
if (!eunomia_is_available()) downloadEunomiaData()
```

In this example we're going to summarise the characteristics of individuals with ankle sprain, ankle fracture, forearm fracture, or hip fracture using the Eunomia synthetic data. 

We'll begin by creating our study cohorts.

```{r}
library(CDMConnector)
library(CodelistGenerator)
library(CohortCharacteristics)
library(dplyr)
library(ggplot2)

con <- DBI::dbConnect(duckdb::duckdb(),
                       dbdir = CDMConnector::eunomia_dir())
cdm <- CDMConnector::cdm_from_con(con,
                                  cdm_schem = "main",
                                  write_schema = "main", 
                                  cdm_name = "Eunomia")

cdm <- generateConceptCohortSet(
  cdm = cdm,
  name = "injuries",
  conceptSet = list("ankle_sprain" = 81151,
                    "ankle_fracture" = 4059173,
                    "forearm_fracture" = 4278672,
                    "hip_fracture" = 4230399),
  end = "event_end_date",
  limit = "all"
)
```

Now we've created our cohorts, we can quickly create a table with an overall summary of the people in each cohort.

```{r}
chars <- cdm$injuries |> 
  summariseCharacteristics()
```

```{r}
tableCharacteristics(chars)
```

```{r}
chars |> 
  filter(variable_name  == "Age") |> 
  plotCharacteristics(plotStyle = "boxplot", 
                      colorVars = "group_level", 
                      xAxis = "group_level", yAxis = "estimate_value", 
                      facet = c("cdm_name"), vertical_x = FALSE)
```

We can also generate summaries that are stratified by some variable of interest. In this case we add an age group variable to our cohort table and then stratify our results by age group.

```{r}
chars <- cdm$injuries |>
  PatientProfiles::addAge(ageGroup = list(c(0,49),
                                          c(50, Inf))) |> 
  summariseCharacteristics(strata = list("age_group"))
```

```{r}
tableCharacteristics(chars,
                     groupColumn = "age_group")
```

```{r}
chars |> 
  filter(variable_name  == "Prior observation") |> 
  plotCharacteristics(plotStyle = "boxplot", 
                      colorVars = "group_level", 
                      yAxis = "group_level", 
                      xAxis = "estimate_value", 
                      facet = c("strata_level"))
```

# Presence in other cohorts

```{r}
meds_cs <- getDrugIngredientCodes(cdm = cdm, 
                                  name = c("acetaminophen",
                                           "morphine",
                                           "warfarin"))
cdm <- generateConceptCohortSet(
  cdm = cdm,
  name = "meds",
  conceptSet = meds_cs,
  end = "event_end_date",
  limit = "all",
  overwrite = TRUE
)
```

```{r}
chars <- cdm$injuries |> 
  summariseCharacteristics(cohortIntersectFlag =  list(
    "Medications prior to index date" = list(targetCohortTable = "meds", 
                         window = c(-Inf, -1)
                             )))
```

These results will automatically be included when we create our table with patient characteristics.

```{r}
tableCharacteristics(chars)
```

We can now also plot our results for these medication cohorts of interset. 

```{r}
chars |> 
  filter(variable_name  == "Medications prior to index date",
         estimate_name == "percentage") |> 
  plotCharacteristics(plotStyle = "barplot", 
                      colorVars = "variable_level", 
                      yAxis = "variable_level", 
                      xAxis = "estimate_value", 
                      facet = c("cdm_name",
                                "group_level"))
```
